\documentclass[journal]{IEEEtran}

\usepackage{newtxtext,newtxmath}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage[colorlinks=true,allcolors=blue]{hyperref}

\graphicspath{{./figures/}}

\begin{document}

\title{Comparative Performance Analysis of Parallel Implementations for Accelerated Streamline Visualization in Vector Field Analysis}
\author{Nate Stott, Rylei Mindrum, Ethan Gee\thanks{6030 High Performance Computing, Utah State University.}}
\markboth{HPC Class Project}{Nate Stott, Rylei Mindrum, Ethan Gee: Accelerated Streamline Visualization}
\IEEEtitleabstractindextext{
	\begin{abstract}
		Streamline visualization is a fundamental tool for analyzing complex vector fields, yet it often faces significant computational overhead due to the intensive numerical integration and irregular data-access patterns required for tracing paths across high-resolution grids.
		This study presents a comprehensive suite of implementations aimed at accelerating streamline visualization through various parallel computing paradigms.
		We developed and evaluated a vector field analyzer across several architectures. Namely, sequential CPU, shared-memory parallelism using Pthreads and OpenMP, distributed-memory systems with OpenMPI, and massively parallel GPU acceleration through CUDA, as well as a hybrid approach utilizing distributed CUDA with OpenMPI.
		Our analysis focuses on the scalability and speedup of each implementation, comparing  performance across various example datasets at a variety of field sizes and streamline complexity.
		Results demonstrate significant performance gains over sequential implementations, with the distributed GPU implementation showing the best performance increase.
	\end{abstract}
	\begin{IEEEkeywords}
		Streamline Visualization, Vector Field Analysis, Parallel Computing, OpenMP, OpenMPI, CUDA, GPU Acceleration, Pthreads.
	\end{IEEEkeywords}
}

\maketitle

\IEEEdisplaynontitleabstractindextext

\IEEEpeerreviewmaketitle

\input{sections/introduction}
\input{sections/background}
\input{sections/methodology}
\input{sections/results}
\input{sections/conclusion}
\input{sections/acknowledgments}
\input{sections/references}

% Add a citation to generate bibliography entries
See \cite{ref1} for more information.

\end{document}
