\section{Conclusion}

This study successfully implemented and evaluated several parallel computing paradigms to accelerate streamline visualization in vector field analysis.
By developing solvers using Pthreads, OpenMP, OpenMPI, and CUDA, we demonstrated the scalability and efficiency of these technologies in handling the computationally intensive and irregular data-access patterns inherent in streamline tracing.

Our findings reveal that OpenMP provides a slight performance advantage over Pthreads in shared-memory environments, primarily due to its superior load-balancing capabilities which were most evident at moderate thread counts.
In GPU-accelerated environments, we observed that tuning block sizes—specifically moving from 128 to 256 threads—significantly impacts occupancy and latency hiding.
Ultimately, the hybrid MPI+CUDA implementation provided the most substantial speedup, proving the effectiveness of combining distributed-memory scaling with massively parallel GPU acceleration.
Crucially, all parallel implementations maintained bitwise parity with the sequential reference, ensuring that performance gains did not come at the cost of numerical accuracy.

Despite these gains, this study was limited to static 1024 $\times$ 1024 grids on the older Kingspeak cluster.
Future research could investigate the performance of these solvers on more modern architectures, such as the A100 or H100 GPUs available on the Notchpeak and Granite clusters.
Additionally, extending the solvers to handle time-varying vector fields and exploring more complex load-balancing strategies for highly non-uniform datasets would further enhance their applicability to large-scale scientific simulations.

