\section{Responsibilities}

\subsection{Nate Stott}
\begin{itemize}
	\item Developed the comprehensive testing and automation pipeline.
	\item Implemented the procedural vector field simulator with HDF5 support.
	\item Integrated OpenMPI and developed the necessary communication wrappers to handle distributed data structures and global synchronization.
	\item Co-authored the CPU-based Disjoint Set Union implementations and optimized for openMPI.
	\item Developed the Python-based visualization tools for qualitative streamline analysis.
	\item Tested the implementations on CHPC.
\end{itemize}

\subsection{Rylei Mindrum}
\begin{itemize}
	\item Architected and implemented both standalone CUDA and hybrid MPI+CUDA streamline solvers.
	\item Designed the parallel Disjoint Set Union (DSU) implementation for the project.
	      as well as implementing in the CUDA implementation to ensure feasibility utilizing atomic synchronization primitives for thread-safe connectivity discovery.
	\item Focused on architectural compliance to ensure code reuse across the C++, CUDA, and OpenMPI implementations.
	\item Tested the implementations on CHPC.

\end{itemize}

\subsection{Ethan Gee}
\begin{itemize}
	\item Developed the project's foundational data structures, including the \texttt{Vec2} representation and the \texttt{FieldGrid} management system.
	\item Implemented the shared-memory parallel solvers using POSIX Threads and OpenMP.
	\item Co-authored the CPU-based Disjoint Set Union implementations and optimized them for shared-memory environments.
	\item Authored the final research report and conducted the comprehensive performance and scaling analysis, including the application of Amdahl's Law to the experimental findings.
\end{itemize}
